### 2020-01-24 authors.tex
Fix things

\newcommand{\sceeiitmandi}{School of Computing and Electrical Engineering, Indian Institute of Technology Mandi, India}

% \author[0000-0000-0000-0000]{The SunPy Community}
% \noaffiliation

% \author[orcid-id]{name}
% \affiliation{}
% below add contributing authors (those folks that have actually worked writing the paper).
% board members get some symbol next to their names.

\author[0000-0002-0870-4665]{Shreyas Bapat}
\altaffiliation{EinsteinPy Lead Developer, EinsteinPy Board Member}
\affiliation{\sceeiitmandi}

\author[0000-0002-1900-3338]{Shilpi Jain}
\altaffiliation{EinsteinPy Board Member}
\affiliation{Department of Earth Sciences, Indian Institute of Technology Roorkee, India}

\author[0000-0002-2369-0628]{Ritwik Saha}
\altaffiliation{EinsteinPy Lead Developer, EinsteinPy Board Member}
\affiliation{\sceeiitmandi}

\author[0000-0002-2784-5666]{Bhavya Bhatt}
\altaffiliation{EinsteinPy Deputy Lead Developer}
\affiliation{\sceeiitmandi}

\author[0000-0003-3354-1893]{Akshita Jain}
\affiliation{\sceeiitmandi}

\collaboration{1000}{(Primary Paper Contributors)}

% Below add contributors to the Sunpy project
% this list is ordered alphabetically

\author[0000-0003-1021-3336]{Priyanshu Khandelwal}
\altaffiliation{EinsteinPy Board Member}
\affiliation{\sceeiitmandi}

\author[0000-0002-1928-1549]{Rishi Sharma}
\altaffiliation{EinsteinPy Board Member}
\affiliation{\sceeiitmandi}

\author[0000-0002-2268-9772]{Sofía Ortín Vela}
\affil{University of Zaragoza, Spain}

\author[0000-0001-8530-7252]{Jialin Ma}
\affil{Georgia Institute of Technology, USA}

\author[0000-0001-7234-6148]{Varun Singh}
\affiliation{\sceeiitmandi}

\author[0000-0002-7179-3795]{Alpesh Jamgade}
\affiliation{Department of Mathematics, Bharath Institute of Higher Education and Research, Chennai, India}

\author[0000-0002-5198-9010]{Manvi Gupta}
\affiliation{\sceeiitmandi}

\author{Tushar Tyagi}
\affiliation{\sceeiitmandi}

\author[0000-0003-4387-8078]{Divya Gupta}
\affiliation{\sceeiitmandi}

\author{Hrishikesh Sudam Sarode}
\affiliation{\sceeiitmandi}

\author[0000-0003-1838-8368]{Tanmay Rustagi}
\affiliation{\sceeiitmandi}

\author[0000-0002-0757-2883]{Abhijeet Manhas}
\affiliation{\sceeiitmandi}

\author[0000-0001-5558-9447]{Raphael Reyna}
\affiliation{California State Polytechnic University, Pomona, USA}

\author[0000-0002-6696-8730]{Govind Dixit}
\affiliation{Information Technology, Indian Institute of Information Technology Lucknow, India}

\author{Ratin Kumar}
\affiliation{National Institute of Technology Kurukshetra, India}

\author[0000-0001-8302-1584]{Sashank Mishra}
\affiliation{Department of Information Technology, Indian Institute of Information Technology Allahabad, India}

\author[0000-0003-0973-4994]{Pratyush Kerhalkar}
\affiliation{Department of Electronics & Communication Engineering, Manipal Institute of Technology, India}

\collaboration{1000}{(EinsteinPy Contributors)}

\correspondingauthor{Shreyas Bapat}
\email{shreyas@einsteinpy.org, bapat.shreyas@gmail.com}